Automated Verification and Synthesis of Embedded Systems: Advances and Challenges

### 2017-03-01 paper.tex
addition until section 4

One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs. In particular, loop invariants, which were computed from programs under verification, must be inductive (and not just invariant) for the corresponding verification conditions to be valid, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}. As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.

There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'', in programming languages with pointers~\cite{pips:2013,Henry:2012}. The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems, using invariant generation and strengthening, mostly based on interval analysis.

Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15,Kinductor,Rocha17}. Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10}, considering the Cell BE processor, and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}. Apart from that, the \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.

Nonetheless, there is still little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in embedded-system verification. That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes. Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.

@@

The basic idea of program synthesis is to automatically construct a program $P$ that satisfies a correctness specification $\sigma$. In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$ as the starting point, and then produce incrementally a sequence of candidate solutions that satisfy $\sigma$~\cite{Abate17}. As a result, a given candidate program $p$ is iteratively refined to match the specification $\sigma$ more closely. Counter-Example Guided Inductive Synthesis (CEGIS) represents one of the most popular approaches to program synthesis currently used in practice~\cite{David15}. Figure~\ref{Counter-Example-Guided-Inductive-Synthesis} shows the basic architecture of the CEGIS approach, which has close connections to algorithmic debugging using counterexamples and abstraction refinement~\cite{Alur13}.

The correctness specification $\sigma$ provided to the program synthesizer is of the form $\exists \vec{F} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{F})$ where $\vec{F}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula typically supported by SMT solvers. The ground terms is interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can be encoded using the SMT solvers bit-vectors theory. The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is updated incrementally. Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$). If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the {\sc verify} phase to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$). If this is the case, then the algorithm terminates; otherwise, additional information is provided to the {\sc synthesize} phase in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again. 
%
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.12]{figure4.eps}
	\caption{Counter-Example Guided Inductive Synthesis.}
	\label{Counter-Example-Guided-Inductive-Synthesis}
\end{figure}

Note that each iteration of the CEGIS loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.  Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times. However, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$. In order to avoid exploring all possible values, machine learning techniques can be used in the {\sc Synthesize} phase to learn from experience, {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.

@@

It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically address programing-language issues and general correctness, without taking into account target applications or system behavior. Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then requires schemes that do not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems. For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.

Indeed, research in software verification is now incorporating such considerations, during checking processes, and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015,Monteiro16} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length (FWL) effects. Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system~\cite{Bessa17}. That ultimately leads to the use of analog-to-digital converters, which are one of the most important aspects to be considered, given that data loss (quantization) is inevitable. As a consequence, the verification procedure has to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations. Additionally, the latter may present a different behavior, regarding what was specified in analog domain, due to the inherent discret-time operation. Indeed, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions, in order to explore the design space. finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

In this respect, DSVerifier has been extended to automatically synthesize digital stabilizing controllers for continuous plants~\cite{Abate17}. In particular, a CEGIS-based approach, implemented in a tool called digital system synthesizer (DSSynth), uses inductive synthesis in conjunction with DSVerifier's algorithm for verifying robust closed-loop stability that addresses plant variations as interval sets as well as FWL uncertainties in digital controllers. DSSynth is able to successfully synthesize stable digital controllers for a set of intricate plant models taken from the control literature within minutes.

Apart from the control system domain, Scratch is another example of a software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA), in the Cell BE processor~\cite{Donaldson10}. That tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races. If support to other DMA operations were added, Scratch could be adapted to different architectures, {\it i.e.}, the same techniques would be employed, but with a different system behavior/knowledge.

@@

State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}. 
%
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{figure3.eps}
	\caption{Verification methodologies for embedded systems.}
	\label{verification-methodologies}
\end{figure}

In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, providing completeness, {\it i.e.},  if all possible states can be reached and evaluated, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness. Additionally, embedded system verification raises additional challenges, such as 
%
\begin{enumerate}
	\item time and energy constraints.
	\item handling of concurrent software.
	\item platform restrictions.
	\item verification of code pieces or modules that rely on larger structures.
	\item legacy designs. %, which are usually developed with low-level languages.
	\item support to different programming languages, frameworks, and interfaces.
	\item Correct code instrumentation.
	\item handling of non-linear and non-convex optimization problems.
\end{enumerate}

Indeed, the first two aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers, and the third one is essential in systems that implement device models, such as digital filters and controllers, which present a behavior that is highly dependent on signal inputs and outputs and whose deployment may be heavily affected by hardware restrictions. The fourth aspect deals with code that uses an existing infrastructure and must be compliant with that, such as Linux kernel modules. The fifth aspect, in turn, is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices. In particular, software developed for that type of embedded systems has been extensively tested and verified, and also optimized for efficiency over years of development. Therefore, when a new product is derived from a given platform, a lot of legacy code is usually reused for improving development time and code quality. The sixth aspect is related to the evolution of development processes and technologies, which may delay the application of suitable verification and synthesis approaches, if verifiers and synthesizers do not support different programming languages and interfaces. Finally, the last one is related to the widespread use of embedded systems in autonomous vehicles navigation systems~\cite{Adouane16}, which demands optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.

Those seven challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where the controlled object ({\it e.g.}, physical plant) typically exhibits continuous behavior whereas the controller (usually implemented by a real-time computer system) operates in discrete time and over a quantized domain. In particular, synthesizers for those systems need to consider the effects of the quantizers (A/D and D/A converters) when a digital equivalent of the controlled object is considered, {\it i.e.}, a model of their physical environment (cf. intelligent product in Figure~\ref{intelligent-product}). Additionally,  finite-precision arithmetic and their related rounding errors need to be considered when correct-by-construction code is generated for embedded systems.

@@

>>>>

One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs. In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to validate the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}. As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.

There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also discover the memory ``shape'', in programming languages with pointers~\cite{pips:2013,Henry:2012}. The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems, using invariant generation and strengthening, mostly based on interval analysis. \textcolor{red}{eh bom citar essa literatura aqui}

Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction and in a completely automatic way ({\it i.e.}, users do not need to provide loop invariants), were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15,Kinductor,Rocha17}. Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10}, considering the Cell BE processor, and (2) do respect time constraints, which are specified during system design phases~\cite{EenS03}. Apart from that, \textit{k}-induction is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.

It is worth noticing that \textit{k}-induction with invariants has the potential to be directly integrated into existing BMC approaches, given that the induction algorithm itself can be seen as an extension after \textit{k} unwindings and it is possible to generate program invariants with other software modules, which are then translated and instrumented into an input program \cite{Rocha15}.

Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in embedded-system verification. That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes. Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.

@@

The basic idea of program synthesis is to automatically construct a program $P$ that satisfies a correctness specification $\sigma$. In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$ as the starting point, and then incrementally produce a sequence of candidate solutions that satisfy $\sigma$~\cite{Abate17}. As a result, a given candidate program $p$ is iteratively refined, in order to match the specification $\sigma$ more closely. Counter-Example Guided Inductive Synthesis (CEGIS) represents one of the most popular approaches to program synthesis that are currently used in practice~\cite{David15}. Figure~\ref{Counter-Example-Guided-Inductive-Synthesis} shows the basic architecture of the CEGIS approach, which has close connections to algorithmic debugging using counterexamples and abstraction refinement~\cite{Alur13}.

The correctness specification $\sigma$ provided to the program synthesizer is of the form $\exists \vec{F} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{F})$, where $\vec{F}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula typically supported by SMT solvers. The ground terms are interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can be encoded using the SMT's bit-vectors part. The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is updated incrementally. Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$, for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$). If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the {\sc verify} phase, in order to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$). If this is the case, then the algorithm terminates; otherwise, additional information is provided to the {\sc synthesize} phase, in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again. 
%
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.12]{figure4.eps}
	\caption{Counter-Example Guided Inductive Synthesis.}
	\label{Counter-Example-Guided-Inductive-Synthesis}
\end{figure}

Note that each iteration of the CEGIS loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.  Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times; however, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$. In order to avoid exploring all possible values, machine learning techniques can be used in the {\sc Synthesize} phase, with the goal of learning from experience, {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.

@@

Currently, SMT-based BMC approaches check code properties in real programs, which basically address programing-language issues and general correctness, without taking into account target applications or system behavior. Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then requires schemes that do not only check if source code is correctly written, but also if it will properly respond in real environments or under external problems. For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.

Indeed, research in software verification is now incorporating such considerations, during checking processes, and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015,Monteiro16} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length (FWL) effects. Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system~\cite{Bessa17}. That ultimately leads to the use of analog-to-digital converters, which are one of the most important aspects to be considered, given that data loss (quantization) is inevitable. As a consequence, the verification procedure has to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations. Additionally, the latter may present a different behavior, regarding what was specified in analog domain, due to the inherent discret-time operation.

In summary, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions, in order to explore the design space. Finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

In this respect, DSVerifier has been extended to automatically synthesize digital stabilizing controllers for continuous plants~\cite{Abate17}. In particular, a CEGIS-based approach, implemented in a tool called digital system synthesizer (DSSynth), uses inductive synthesis in conjunction with DSVerifier's algorithm for verifying robust closed-loop stability that addresses plant variations as interval sets, as well as FWL uncertainties in digital controllers. DSSynth is able to successfully synthesize stable digital controllers for a set of intricate plant models, taken from the control literature, within minutes.

Apart from the control system domain, Scratch is another example software model-checking tool, which uses knowledge about the system to be verified and the underlying hardware, with the goal of detecting races related to direct memory access (DMA), in the Cell BE processor~\cite{Donaldson10}. That tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races. If support to other DMA operations were added, Scratch could be adapted to different architectures, {\it i.e.}, the same techniques would be employed, but with a different system behavior/knowledge.

One may also notice that such a paradigm, {\it i.e.}, verification and synthesis based on an expected system behavior, is not restricted to digital filters and controllers or DMA races, but it can also by applied to possibly any real system, as long as the desired behavior can be expressed as properties in BMC frameworks. For instance, regarding self-driving cars, a property could be the detection of pedestrians or animals, or even bicycles. Indeed, the latter is considered one of the most difficult problems, due to the myriad of possible shape and colors \cite{selfcar}. In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised \cite{BMCml}.

@@

Generally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}. 
%
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{figure3.eps}
	\caption{Verification methodologies for embedded systems.}
	\label{verification-methodologies}
\end{figure}

In particular, the main challenges regarding verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, providing completeness, {\it i.e.},  if all possible states can be reached and evaluated, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness. Additionally, embedded system verification raises additional challenges, such as 
%
\begin{enumerate}
	\item time and energy constraints.
	\item handling of concurrent software.
	\item platform restrictions.
	\item verification of code pieces or modules that rely on larger structures.
	\item legacy designs. %, which are usually developed with low-level languages.
	\item support to different programming languages, frameworks, and interfaces.
	\item correct code instrumentation.
	\item handling of non-linear and non-convex optimization problems.
\end{enumerate}

Indeed, the first two aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers, and the third one is essential in systems that implement device models, such as digital filters and controllers, which present a behavior that is highly dependent on signal inputs and outputs and whose deployment may be heavily affected by hardware restrictions. The fourth aspect deals with code that relies on existing infrastructures and must be compliant with those, such as Linux kernel modules. The fifth aspect, in turn, is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices. In particular, software developed for that type of embedded systems has been extensively tested and verified, and also optimized for efficiency over years of development. Therefore, when a new product is derived from a given platform, a lot of legacy code is usually reused for improving development time and code quality. The sixth aspect is related to the evolution of development processes and technologies, which may delay the application of suitable verification and synthesis approaches, if verifiers and synthesizers do not support different programming languages and interfaces. The seventh aspect highlights a subjective task in verification: where one must insert an evaluation point or assertion, in order to correct address a property. Finally, the last one is related to the widespread use of embedded systems in autonomous-vehicle navigation systems~\cite{Adouane16}, which demands optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.

Those eight challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where the controlled object ({\it e.g.}, physical plant) typically exhibits continuous behavior, whereas the controller (usually implemented by a real-time computer system) operates in discrete time and over a quantized domain. In particular, synthesizers for those systems need to consider the effects of quantizers (A/D and D/A converters), when a digital equivalent of the controlled object is considered, {\it i.e.}, a model of their physical environment (cf. intelligent product in Figure~\ref{intelligent-product}). Additionally,  finite-precision arithmetic and their related rounding errors need to be considered, when correct-by-construction code is generated for embedded systems.

@@

\bibitem{selfcar}
Fairley P.:
\newblock {\em {Self-Driving Cars Have a Bicycle Problem}}.
\newblock http://spectrum.ieee.org/transportation/self-driving/selfdriving-cars-have-a-bicycle-problem.

\bibitem{BMCml}
L.~Bortolussi, D. Milios, G, Sanguinetti.:
\newblock {\em {Smoothed model checking for uncertain Continuous-Time Markov Chains}}.
\newblock In:. Information and Computation, 2016.
